---
title: "Hot Dogs and Pizza in Baltimore"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Downloading Data

  First of all, download the file to your pc

```{r}
vendors <- "https://data.baltimorecity.gov/api/views/bqw3-z52q/rows.csv?accessType=DOWNLOAD"
dir.create("./TempData")
download.file(vendors,destfile = "./TempData/BFood.csv")
if (file.exists("./TempData/BFood.csv"))
{
  tam <- file.info("./TempData/BFood.csv")$size
  paste("File downloaded, ",tam," bytes")
} else
{
  "Error downloading file!"
}
```

###Importing CSV Table
  
  Now import the csv file to R environment (DataFrame Format).

```{r}
bVendors <- read.csv(file="./TempData/BFood.csv", header=TRUE, sep=",", stringsAsFactors=FALSE)
str(bVendors)
```

###Eliminating unusable variables

  If you check bVendors DataFrame, there will be some unmeaningful variables like bVendors$Id. So you can remove it.

```{r}  
#Turning LicenseNum to Factor
bVendors$LicenseNum <- as.factor(bVendors$LicenseNum)

bVendors$Id
bVendors$Id <- NULL
bVendors$St <- NULL
  
```

###Changing variable names
  
  Location.1 is not a good name for a variable, let's change it to just "location".
  
```{r} 

names(bVendors)[names(bVendors) == "Location.1"] <- "location"
str(bVendors)
   
```

###Extracting Latitude and Longitude from laction

  Latitude and Longitude is a very important data, it should be in it's own variable.
  
  Let's extract it from location using Regular Expression.
  
```{r} 
#regex match for coordinates
tempS <- regmatches(bVendors$location,gregexpr("[0-9.]+",bVendors$location))
lats <- vector(length = nrow(bVendors),mode = "numeric")
longs <- vector(length = nrow(bVendors),mode = "numeric")

#for each row, extract the lat and long
for(i in 1:nrow(bVendors)) 
{
  lats[i] <- as.numeric(tempS[[i]][2])
  longs[i] <- as.numeric(tempS[[i]][3])
}

#adding to bVendors Data Set
bVendors$lat <- lats
bVendors$long <- longs
str(bVendors)

head(subset(bVendors, select = c(location,lat,long)))
  
```

### Hot Dogs!

  What can i do to check which vendors sells hot dogs in baltimore?
  
  You can check in the variable ItemsSold if the vendor have Hot dog, using grepl
  
```{r}
bVendors$hotdog <- grepl("Hot dog",bVendors$ItemsSold)
head(subset(bVendors, select = c(ItemsSold,hotdog)))
```

  Ok... 
  The vendor sells "Hot Dogs".
  How can we ensure that all variations of "hot dogs" are represented?
  
  Turning ItemsSold to lower case and looking for "hot dog"!
  
  
```{r}
bVendors$ItemsSold = tolower(bVendors$ItemsSold)


bVendors$hotdog <- grepl("hot dog",bVendors$ItemsSold)
head(subset(bVendors, select = c(ItemsSold,hotdog)))
```

### If you preffer pizza, then...

Just create the pizza variable

```{r}
bVendors$pizza <- grepl("pizza",bVendors$ItemsSold)
head(subset(bVendors, select = c(ItemsSold,pizza)))

```

###Getting city names

using string split with regex, you can get the city names

```{r}
city = c()

for(i in 1:nrow(bVendors)) 
{
  city[i] <- strsplit(bVendors$location[i], split = "(?<=[a-zA-Z])\\s*(?=[0-9])", perl = TRUE)[[1]][1]
}
  city = unlist(city)
  
  bVendors$city = city
  
  bVendors$city
```
