---
title: "Climatic and Geological Zones"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Libraries

  sqldf is a library to use R data.frames as SQL table.
  
```{r}
library(sqldf)
```


###Importing covtype.csv and soiltype.csv

  Suposing that covtype.csv is in current directory

```{r}
    covtype <- read.csv("covTypev1.csv", sep=",", stringsAsFactors=FALSE)
```

  Using a temporary variable to hold SoilType codes 

```{r}
code <- covtype$SoilType
code <- gsub("ST0", "", code)
code <- gsub("ST", "", code)

soiltype <- data.frame(code)

```

  SoilType informations can be found in the information file of covtype data
  
  https://archive.ics.uci.edu/ml/machine-learning-databases/covtype/covtype.info
  
  
  Suposing that information data was already extracted from information file, import  SoilType file
  
  

```{r}

  soilindex <- read.csv("soiltype.csv", header=FALSE, sep=",", stringsAsFactors=FALSE)

``` 

### Selecting and manipulating data

Extracting data from SoilType to the main Data Frame (covtype)

```{r}

  #Joining ELU code to covtype data.frame using sqldf
  elu <- sqldf('select soilindex.V2 from soiltype left outer join soilindex on soiltype.code = soilindex.V1')
  
  covtype$elu <- unlist(lapply(elu$V2, as.character))
  str(covtype)
  
``` 

Extracting codes to Climatic Zones and Geological Zones and Importing Data From information file

```{r}

climazone <- substring(covtype$elu, first =1, last = 1)
geozone <- substring(covtype$elu, first =2, last = 2)

#adding to covtype data.frame
covtype$clicode <- unlist(climazone)
covtype$geocode <- unlist(geozone)


climazone <- read.csv("climazone.csv", header=FALSE, sep=",", stringsAsFactors=FALSE)
geozone <- read.csv("geozone.csv", header=FALSE, sep=",", stringsAsFactors=FALSE)

``` 

Selecting Data to the main Data Frame

```{r}
climazone <- sqldf('select climazone.V2 from covtype left outer join climazone on covtype.clicode = climazone.V1')
geozone <- sqldf('select geozone.V2 from covtype left outer join geozone on covtype.geocode = geozone.V1')

covtype$clizone <- unlist(climazone)
covtype$geozone <- unlist(geozone)

head(covtype)
``` 

Done!


